Grove: A Convergent Collaborative Structure-Editor Calculus

### 2020-07-14 grove-paper.tex
Updates from Overleaf

We model edits to the code as sequences of deletions and additions of edges to this graph.
For example, consider the first graph in Figure~\ref{fig:Delete and Add}, which
represents the expression \texttt{1 + 2}.
Suppose we want to change the \texttt{2} into a \texttt{3} so the expression
becomes \textt{1 + 3}.
We model this by first deleting edge~7.
This results in the middle graph in Figure~\ref{fig:Delete and Add}, which
represents the expression \texttt{1 + \hole}.
Note that vertex~9 still exists and can be manipulated (which we will exploit in the next subsection).
Vertex~9 is just not reachable from the root.

@@

sNote that we will omit orphans from diagrams if they are not relevant
(e.g., they are deleted and are not added back later)

\subsubsection{Move}
\begin{figure}
\centering
\begin{tikzpicture}
\path (0,0)    node {\includegraphics[scale=0.75]{graphviz/move/1.dot}};
\path (0,-5cm) node {\includegraphics[scale=0.75]{graphviz/move/2.dot}};
\path (0,-9cm) node {\includegraphics[scale=0.75]{graphviz/move/3.dot}};
\end{tikzpicture}
\caption{Move}
\label{fig:Move}
\end{figure}

\subsubsection{Cut and Paste}

cut/paste (i.e., move) = delete then add
copy structure vs copy id
paste structure vs paste id
  (show two or three options at paste time)

@@

@@

>>>>

We model edits to the code as sequences of deletions and additions of edges to this graph.
For example, consider the first graph in Figure~\ref{fig:Delete and Add}, which
represents the expression \texttt{1 + 2}.
Suppose we want to change the \texttt{2} into a \texttt{3} so the expression
becomes \texttt{1 + 3}.
We model this by first deleting edge~7.
This results in the middle graph in Figure~\ref{fig:Delete and Add}, which
represents the expression \texttt{1 + \hole}.
Note that vertex~9 still exists and can be manipulated (which we will exploit in the next subsection).
Vertex~9 is just not reachable from the root.

@@

Note that we will omit orphans from diagrams if they are not relevant
(e.g., they are deleted and are not added back later)

\subsubsection{Move}

\begin{figure}
\centering
\begin{tikzpicture}
\path (0,0)    node {\includegraphics[scale=0.75]{graphviz/move/1.dot}};
\path (0,-5cm) node {\includegraphics[scale=0.75]{graphviz/move/2.dot}};
\path (0,-9cm) node {\includegraphics[scale=0.75]{graphviz/move/3.dot}};
\end{tikzpicture}
\caption{Move}
\label{fig:Move}
\end{figure}

An interesting consequence of our design is that a move operation
does not have to delete code from one position and create a copy
in another position.
Instead, by deleting just the edge to the code to be moved and then
adding a new edge to that code somewhere else, the actual code is
moved, not just a copy.

For example, consider Figure~\ref{fig:Move}, which represents
the movement of \texttt{1 * 2} from the left child of \texttt{plus}
to its right child.  The top graph is the code before the move.
First, edge~7 is deleted, which results in the middle graph
where the rooted code is now \texttt{\hole{} + \hole},
and vertex~6 is orphaned.

Next edge~16 is added, which results in the bottom graph
where the rooted code is now \texttt{\hole{} + 1 * 2},
and vertex~6 is no longer orphaned.

This move semantics gives us a richer structure than
when treating code as a list of lines.  We exploit
this in Section~REF:TODO in order merge edits that involve moving code.

Of course, if the user wants a genuine copy
and not just a move, that can easily be achieved
(TODO: figure for this?)

This is similar to the paste with formatting versus
paste without formatting.

@@

When editing different parts of the code

For example, consider the graphs in Figure~\ref{fig:Editing Different Parts of the Code}.
The

@@

(Similar to how git merged across moves.
A notable difference though is that \texttt{git} detects
file moves by structural similarity rather than
our nominal equality.
(We exploit this in Subsection~REF:TODO).
